$\pi$SOD-M: A Methodology for Building Service-Oriented Applications in the Presence of Non-Functional Properties

### 2013-12-12 JournalToSubmit2013/modelingFlyingPig.tex
Value model and BPMN

Figure~\ref{fig:E3valuemodel} shows the E3 value model~\cite{e3value} for \textsl{FlyingPig}.
The E3 value model is a business model that represents a business case graphically as a set of value exchanges ($\bigtriangleup\ \bigtriangledown$) and value activities (rounded boxes) performed by business actors (squared boxes).

In our use-case, we identify two business actors: ORCA and the client company that interact to produce value.
Value exchanges are determined by a service (provided by ORCA) and information and economic values (provided by the client).

{\color{magenta} Valeria: Could you please complete these ideas? --G \& M.}

Figure~\ref{fig:bpmn} shows the BPMN model\footnote{Details on BPMN (Business Process Management Notation) can be found in http://www.bpmn.org/} for the scenario. 
There are three business actors: ORCA, the Client company and a User.
The two latter business actors belong to a channel, which is a third-party broker that manages the communication between GCP Global and the Client company. 
The User is a contact member of the Company, who will coordinate the assessment process.
This process will involve other members of the company as well.

The process  starts by a request from a company branch asking for an assessment. 
The request leads to the definition of a group of users that will answer questionnaires for evaluating risk. 
Questionnaires are considered \textit{tasks} that users will have to perform. 
Given a list of tasks a user decides whether she takes on the task (i.e., answering a questionnaire) or delegate it to another user.

During the process of answering a questionnaire, the answer of a question can trigger the generation of other questionnaires, that become new tasks to be executed. Once a questionnaire has been completely answered it is stored and processed to generate a set of un-compliant situations associated to their corresponding amendments if there are any or a report specifying a compliance level, incidents and a risk map.

{\color{magenta} Valeria and Khalid: Could you please complete these ideas? --G \& M.}

>>>>

Figure~\ref{fig:E3valuemodel} shows the value model~\cite{e3value} for the case of \textsl{FlyingPig}.
The value model is a business model that represents a business case graphically as a set of value exchanges ($\bigtriangleup\ \bigtriangledown$) and value activities (rounded boxes) performed by business actors (squared boxes).

In our use-case, we identify two business actors: \textsl{ORCA} and \textsl{Broker}. Brokers are the responsible for channeling requests for risk assessment of one or several companies. ORCA have two value activities which are services that provide it an economical benefit: to \textsl{identify amendments} and the possibility to \textsl{asses risk situation}. The value exchanged between ORCA and Brokers are: \textsl{amendments} and \textsl{evaluation reports} which are value objects ([...]) for the any companies that need to have a risk assessment, and the \textsl{questionnaire and evidences} and the risk assessment \textsl{fee} which are value objects for ORCA System.

Beside the static representation of a business case, the value model allows for the representation of dependency paths which enhance the understanding of a business idea by showing all value exchanges triggered by the occurrence of an end-consumer need. A dependency path has a direction and consists of dependency nodes linked by mean of connections. A dependency node is a start stimulus, an AND-fork or AND-join, an OR-fork or OR-join, or an end node (see Legend on Figure~\ref{fig:E3valuemodel}). An start stimulus represents a trigger for the exchange of economic value objects while an end node represents a model boundary.

Figure~\ref{fig:E3valuemodel} shows the dependency paths for \textsl{FlyingPig}. The dependency path identified in that case is the need of assessment of a particular company. It denotes a stimulus caused by the client company who ask for risk assessment. Once this need occurs the value exchanges between ORCA and Broker are triggered. The value model represent that ORCA will provide amendments and an evaluation report in return for the questionnaire and evidences and the fees provided by the Broker.

{\color{blue} Javier: Please modify the value model to include fee as a value object in return for evaluation report. Modified figure is on the git}

A better understanding of the process in which these value exchanges occurs may be modeled using a BPMN model. Figure~\ref{fig:bpmn} shows the BPMN model\footnote{Details on BPMN (Business Process Management Notation) can be found in http://www.bpmn.org/} for the scenario. It includes two pools representing the \textsl{ORCA} system and the \textsl{Brokers}. Brokers have two lanes, the client \textsl{Company} and a \textsl{User}. The user is a contact member of the company, who will coordinate the assessment process. This process will involve other members of the company as well.

The process starts by a request from a company asking for a risk assessment. Note that it is in agreement with the explanation of the start stimulus on the value model above. The request leads to the definition of a group of users that will answer questionnaires for evaluating risk. Questionnaires are considered tasks that users will have to perform. Given a list of tasks, the company assign a responsible and delegate the task. A task may be to answer a questionnaire or add an evidence but also could implies to solve some "risky situation" (e.g. to repair a door {\color{blue} ...¿Genoveva mira por favor si esto puede ser o hay algun ejemplo mas bueno/serio???} ).

Once a tasks are completed they are stored and analyzed to generate a set of un-compliant situations associated to their corresponding amendments if there are any or a report specifying a compliance level, incidents and a risk map.
During the process of analysing a questionnaire, the answers of some questions can trigger the generation of other questionnaires or amendments, that become new tasks to be executed.
